HUME: Human UCCA-based Machine-Translation Evaluation

### 2016-03-14 paper/main.tex
some comments and corrections on 1-3

\newcommand{\com}[1]{}
\newcommand{\secref}[1]{Section \ref{#1}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newcommand{\tabref}[1]{Table~\ref{#1}}
\newcommand{\oa}[1]{\footnote{\color{red} #1}}
\newcommand{\bh}[2]{\footnote{\color{blue} #1}}

@@

  Recent interest in semantics-aware MT systems underscores the importance of
  semantic evaluation for Machine Translation (MT) systems. 
  A semantic measure, sensitive to possible semantic discrepencies
  between the translation output and the source, is likely support the construction
  and tuning of structure-aware MT, much in the same way that string-based measures,
  such as BLEU, enabled progress in shallower MT.
  We present a novel human semantic evaluation measure for MT, Human
  UCCA-based MT evaluation (HUME), building on the UCCA semantic representation scheme.
  Making progress over previous work, the proposed measure takes into account
  a wider range of semantic phenomena and does not rely on semantic annotation
  of the MT output itself, whose interpretation is often unclear.
  We experiment with four language pairs, demonstrating HUME's broad applicability,
  as well as its reliability, reflected in its inter-annotator agreement rates.

@@

These problems were partially addressed by measures that decompose over sub-parts of the evaluated
translation sentence (henceforth, {\it translation}),
most commonly over its words or n-grams,
quantifying the overlap of its sub-parts relative a reference translation.
More recent work on MT evaluation proposed measures
that decompose over semantically defined units,
quantifying the similarity of the output and the reference in terms of
their verb argument structure (the most notable of these measures is
MEANT \cite{lo2011structured}).
While allowing to localize semantic errors in the translation,
these measures' focus on verb argument structures misses out on many pervasive semantic phenomena,
such as inter-clause linkage and nominal argument structures, and thus only provides
a limited perspective on the semantic similarity between the output and reference.
Moreover, assigning semantic structure to MT output, especially when low in quality,
is difficult and arguably ill-defined.

@@

We conduct experiments with four language pairs: English-German, English-Polish,
English-Romanian and English-Czech, and report efficiency (time of annotation) and
inter-annotator agreement scores, as well as qualitative comparison of the resulting
HUME scores with crowd-sourced adequacy judgements. Results show that HUME obtains 
agreement scores which are on par with those of standard sentence-level ranking measures,
but unlike them, HUME's scores do not decrease with logner sentences.
The paper concludes by outlining ongoing work to construct semi- and
fully-automatic variants of HUME.

@@

\paragraph{Machine Translation Evaluation.}
%MT evaluation has been traditionally pursued through human-based
%evaluation that quantifies the adequacy of the translation in preserving the meaning 
%of the source sentence, and its fluency as a sentence in the target language \cite{lopez2008statistical}.
Human evaluation is generally determined
either according to the relative ranking of the outputs of multiple systems
e.g., in the WMT shared tasks \cite{bojar2015findings}, or by assigning
indiviaul adequacy/fluency scores to each translation, a procedure recently improved
by \newcite{graham2015improving}.
However, while providing the gold standard for MT evaluation, human evaluation is not a scalable solution,
and is unfeasible where a wide variety of systems or parameter settings are to be evaluated.
% has motivated the construction of cheaper methods that correlate with human
%judgement.

@@

\paragraph{Semantic Representation.}
UCCA (Universal Conceptual Cognitive Annotation) is a cross-linguistically applicable, lightweight
scheme for semantic annotation. Formally, a UCCA structure is a directed acyclic graph (DAG),
whose leaves correspond to the words of the text.
The graph’s nodes, called {\sc units}, are either terminals or several elements jointly
viewed as a single entity according to some semantic or cognitive consideration. Edges bear
a category, indicating the role of the sub-unit in the relation the unit represents. UCCA
currently focuses on annotating argument structures (adjectival, nominal, verbal and others) and
relations between them. The most basic notion is the {\sc scene}, which describes a movement, an
action or a state which is persistent in time. Each Scene contains one main relation, as well
as one or more participants. For example, the sentence ``After graduating, Tom moved to NYC''
contains two Scenes, whose main relations are ``graduating'' and ``John''. The participant ``John''
is a part of both Scenes, while ``NYC'' only of the latter. Further categories account for
relations between Scenes and the internal structures of complex participants and relations.
See \figref{fig:ucca_example}.

@@

There are two manual steps in the annotation process. The first is to create UCCA annotations for the source sentence,
following the
UCCA guidelines\footnote{All UCCA-related resources can be found here: \url{http://www.cs.huji.ac.il/~oabend/ucca.html}}. Source sentences are all in English, and the annotation was performed by four
computational linguists\oa{were there no other annotators from Prague?}. After we have the UCCA annotation on 
the source, we use the word-to-word alignments from the MT system in order to project this annotation to the MT output,
and use native speakers of the target language in order to annotate this output. Since these annotators have to
judge the quality of the translation, they also needed to be fluent in English.\bh{Somewhere we could explain that we
could do a similar annotation with monolinguals, although it entails aligning reference and MT output}

The basic strategy for annotating HUME on the  MT output to step through the source sentence semantic 
components (projected UCCA nodes), looking at the translation via the word alignments, and marking
on the source structure which parts have been correctly translated.
There are two kinds of components (nodes): a word or basic semantic unit, and a structural component
which contains one or more sub-components.

>>>>

\newcommand{\com}[1]{}
\newcommand{\secref}[1]{Section \ref{#1}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newcommand{\tabref}[1]{Table~\ref{#1}}
\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
\newcommand{\bh}[2]{\footnote{\color{blue}BH: #1}}

@@

  Recent interest in semantics-aware MT systems underscores the importance of
  semantic evaluation for Machine Translation (MT) systems. 
  A semantic measure, sensitive to possible semantic discrepencies
  between the translation output and the source, is likely support the construction
  and tuning of structure-aware\bh{What is ``structure-aware MT''} MT, much in the same way that string-based measures,
  such as BLEU, enabled progress in shallower MT.
  We present a novel human semantic evaluation measure for MT, Human
  UCCA-based MT evaluation (HUME), building on the UCCA semantic representation scheme.
  Making progress over previous work, the proposed measure takes into account
  a wider range of semantic phenomena and does not rely on semantic annotation
  of the MT output itself, whose interpretation is often unclear.
  We experiment with four language pairs, demonstrating HUME's broad applicability,
  as well as its reliability, reflected in its inter-annotator agreement rates.

@@

These problems were partially addressed by measures that decompose over sub-parts of the evaluated
translation sentence (henceforth, {\it translation}),
most commonly over its words or n-grams,
quantifying the overlap of its sub-parts relative a reference translation.
More recent work on MT evaluation proposed measures
that decompose over semantically defined units,
quantifying the similarity of the output and the reference in terms of
their verb argument structure (the most notable of these measures is
MEANT \cite{lo2011structured})\bh{Should we just refer to HMEANT, as that is the comparable measure from the MEANT family?}.
While allowing to localize semantic errors in the translation,
these measures' focus on verb argument structures misses out on many pervasive semantic phenomena,
such as inter-clause linkage and nominal argument structures, and thus only provides
a limited perspective on the semantic similarity between the output and reference.
Moreover, assigning semantic structure to MT output, especially when low in quality,
is difficult and arguably ill-defined.

@@

We conduct experiments with four language pairs: English-German, English-Polish,
English-Romanian and English-Czech, and report efficiency (time of annotation) and
inter-annotator agreement scores, as well as qualitative comparison of the resulting
HUME scores with crowd-sourced adequacy judgements. Results show that HUME obtains 
agreement scores which are on par with those of standard sentence-level ranking measures,
but unlike them, HUME's scores do not decrease with longer sentences.\bh{We may have to focus
more on qualitative aspects, if we just try to sell the measure on IAA, then the figures
are not stunning.}
The paper concludes by outlining ongoing work to construct semi- and
fully-automatic variants of HUME.

@@

\paragraph{Machine Translation Evaluation.}
%MT evaluation has been traditionally pursued through human-based
%evaluation that quantifies the adequacy of the translation in preserving the meaning 
%of the source sentence, and its fluency as a sentence in the target language \cite{lopez2008statistical}.
Human evaluation is generally determined
either according to the relative ranking of the outputs of multiple systems
e.g., in the WMT shared tasks \cite{bojar2015findings}, or by assigning
individual adequacy/fluency scores to each translation, a procedure recently improved
by \newcite{graham2015improving}.
However, while providing the gold standard for MT evaluation, human evaluation is not a scalable solution,
and is unfeasible where a wide variety of systems or parameter settings are to be evaluated.
% has motivated the construction of cheaper methods that correlate with human
%judgement.

@@

\paragraph{Semantic Representation.}
UCCA (Universal Conceptual Cognitive Annotation) is a cross-linguistically applicable, lightweight
scheme for semantic annotation. Formally, a UCCA structure is a directed acyclic graph (DAG),
whose leaves correspond to the words of the text.
The graph’s nodes, called {\sc units}, are either terminals or several elements jointly
viewed as a single entity according to some semantic or cognitive consideration. Edges bear
a category, indicating the role of the sub-unit in the relation the unit represents. UCCA
currently focuses on annotating argument structures (adjectival, nominal, verbal and others) and
relations between them. The most basic notion is the {\sc scene}, which describes a movement, an
action or a state which is persistent in time. Each Scene contains one main relation, as well
as one or more participants. For example, the sentence ``After graduating, Tom moved to NYC''
contains two Scenes, whose main relations are ``graduating'' and ``John''\bh{Don't you mean ``graduating'' and
``moved''? And where did ``John'' come from?}. The participant ``John''
is a part of both Scenes, while ``NYC'' only of the latter. Further categories account for
relations between Scenes and the internal structures of complex participants and relations.
See \figref{fig:ucca_example}.

@@

There are two manual steps in the annotation process. The first is to create UCCA annotations for the source sentence,
following the
UCCA guidelines\footnote{All UCCA-related resources can be found here: \url{http://www.cs.huji.ac.il/~oabend/ucca.html}}. Source sentences are all in English, and the annotation was performed by four
computational linguists\oa{were there no other annotators from Prague?}. After we have the UCCA annotation on 
the source, we use the word-to-word alignments from the MT system in order to project this annotation to the MT output,
and use native speakers of the target language in order to annotate this output.\bh{Is it obvious how this was done? What about $n$-$m$ alignments?} Since these annotators have to
judge the quality of the translation, they also needed to be fluent in English.\bh{Somewhere we could explain that we
could do a similar annotation with monolinguals, although it entails aligning reference and MT output}

The basic strategy for annotating HUME on the  MT output to step through the source sentence semantic 
components (projected UCCA nodes), looking at the translation via the word alignments, and marking
on the source structure which parts have been correctly translated. A typical display from the HUME annotation tool is
shown in \figref{mttool}, where we see the translation mapped onto UCCA nodes and corresponding source words.

There are two kinds of nodes: a word or basic semantic unit, and a structural component
which contains one or more sub-components.
